---
title: "p8105_hw2_rw2844"
author: "Renjie Wei"
date: "2020/9/24"
output: 
  github_document:
    toc: TRUE
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE,message=FALSE,tidy=TRUE)
```

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
```
# Problem 1

Read and create `mr_trash` dataset.
```{r import_data, warning=FALSE}
trash_path = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx"
mr_trash = 
  read_excel(
  path = trash_path,
  sheet = "Mr. Trash Wheel",
  range = cell_cols("A:N")) %>%
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read and clean the precipitation data.
```{r precip_data}
precip_18 =
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2018 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2018) %>% 
	relocate(year) 

precip_17 = 
	read_excel(
		"./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2017 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2017) %>% 
	relocate(year)
```

Combine two datasets.
```{r merge_data}
month_df = 
	tibble(
		month = 1:12,
		month_name = month.name
	)
precip_df =
	bind_rows(precip_18, precip_17)
precip_df = 
	left_join(precip_df, month_df, by = "month")
head(precip_df)
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(mr_trash)` rows in our final dataset. Additional data sheets include month precipitation data. 

* In this dataset:

  + The median number of sports balls found in a dumpster in 2017 was `r mr_trash %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`
  
  + The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.

# Problem 2

Read and clean the NYC Transit data.

```{r nyc_data}
nyc_trans_df = 
  read_csv(
    "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
  ) %>%
  janitor::clean_names() %>% 
  select(
    .,line, station_name, station_latitude, station_longitude, starts_with("route"), entry, vending, entrance_type, ada
    ) %>% 
  mutate(
    entry = recode(entry, "YES" = TRUE, "NO" = FALSE),
    vending = recode(vending, "YES" = TRUE, "NO" = FALSE),
      )
  
#check whether the variables are right or not
str(nyc_trans_df)

```

In the chunk above, we loaded the NYC Transit dataset, cleaned the names of columns, selected the desired variables and changed two character variables to logical variables ---- `entry` and `vending`. There are `r nrow(nyc_trans_df)` rows and `r ncol(nyc_trans_df)` columns in the `nyc_trans_df`. The dataset contained the following variables: `r colnames(nyc_trans_df)`. These data are not tidy.


```{r using_these_data, include = FALSE}
station_distinct =
  distinct(
    nyc_trans_df, line, station_name, .keep_all = TRUE
    )

ada_compliant = 
  filter(
    station_distinct, ada == TRUE
  )

```

```{r using_data_prop, include = FALSE}
n1 = 
  filter(
    nyc_trans_df, vending == FALSE & entry == TRUE
  ) %>% 
  nrow()

n2 = 
  filter(
    nyc_trans_df, vending == FALSE
  ) %>% 
  nrow()

```


* **Answers:**
  + There are `r nrow(station_distinct)` stations in NYC.
  + And `r nrow(ada_compliant)` of them are ADA compliant.
  + The proportion of station entrances / exits without vending allow entrance is `r n1/n2`.

Reformat the data:
```{r reformat_data}
# reformat
dist_route_df = station_distinct %>%
  # convert
  mutate_at(vars(route8:route11), as.character) %>% 
  pivot_longer(
    route1:route11,
    names_to = "route",
    values_to = "train")

head(dist_route_df)
```

 + The number of distinct stations serve the A train is `r nrow(filter(dist_route_df, train == "A"))`.
 + The number of distinct stations serve the A train and are ADA compliant is `r nrow(filter(dist_route_df, train == "A" & ada == "TRUE"))`.

# Problem 3

Read and clean the pols-month dataset:
```{r pols_month}
pols_df =
  read_csv("./data/fivethirtyeight_datasets/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  #separate mon
  separate(
    mon, 
    c("year", "month", "day"), 
    convert = TRUE
    ) %>%
  mutate(month = month.name[month]) %>%
  mutate(
    president = recode(prez_gop, '1' = "gop", '0' = "dem")
    ) %>%
  # remove and organize variables
  select(-prez_dem, -prez_gop, -day) %>%
  relocate(year, month, president)

head(pols_df)
```

Read and clean the snp dataset:
```{r snp}
snp_df =
  read_csv("./data/fivethirtyeight_datasets/fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  separate(
    date, 
    c("month", "day", "year"), 
    convert = TRUE
    ) %>%
  mutate(month = month.name[month]) %>%
  select(-day) %>%
  relocate(year, month)

head(snp_df)
```

Read and clean the unemployment dataset:
```{r unemp}
unemp_df =
  read_csv("./data/fivethirtyeight_datasets/fivethirtyeight_datasets/unemployment.csv") %>%
  #wide -> long
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment_rate"
  ) %>%
  mutate(month = match(month, month.abb)) %>%
  mutate(month = month.name[month]) %>%
  janitor::clean_names() %>%
  mutate(year = as.integer(year))

head(unemp_df)
```
Merge them!
```{r merge}
pols_snp = left_join(
  pols_df, snp_df, by = c("year", "month")
  )
merge_df = 
  left_join(
    pols_snp, unemp_df, by = c("year", "month")
    ) %>%
  relocate(year, month, president, close, unemployment_rate)


head(merge_df)
```


* In the `pols_df`, there are`r nrow(pols_df)`rows and `r ncol(pols_df)` columns. The year ranges is `r min(pull(pols_df, year))` to `r max(pull(pols_df, year))`. The variables are: `r colnames(pols_df)`.
  + `mon`: date of the count
  + `prez_gop`: indicator of whether the president was republican on the associated date (1 = yes, 0 = no)
  + `gov_gop`: the number of republican governors on the associated date
  + `sen_gop`: the number of republican senators on the associated date
  + `rep_gop`: the number of republican representatives on the associated date
  + `prez_dem`: indicator of whether the president was democratic on the associated date (1 = yes, 0 = no)
  + `gov_dem`: the number of democratic governors on the associated date
  + `sen_dem`: the number of democratic senators on the associated date
  + `rep_dem`: the number of democratic representatives on the associated date


* For `snp_df`,there are`r nrow(snp_df)`rows and `r ncol(snp_df)` columns. The year ranges is `r min(pull(snp_df, year))` to `r max(pull(snp_df, year))`. The variables are: `r colnames(snp_df)`.
  + `close`:the closing values of the S&P stock index on the associated date


* For `unemp_df`,there are`r nrow(unemp_df)`rows and `r ncol(unemp_df)` columns. The year ranges is `r min(pull(unemp_df, year))` to `r max(pull(unemp_df, year))`. The variables are: `r colnames(unemp_df)`.
  + `Jan:Dec`:percentage of unemployment in January-December of the associated year


* For the `merge_df`, there are`r nrow(merge_df)`rows and `r ncol(merge_df)` columns. The year ranges is `r min(pull(merge_df, year))` to `r max(pull(merge_df, year))`.
  + This merged dataset tells us the number of national politicians, the closing value, and the unemployment rate. 